---
title: "Practical_Machine_Learning_Project"
author: "Dalton Sloan"
output: html_document
---

###Assignment
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 

###Data
The training data for this project are available [here](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv) 

The test data are available [here](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv)

The data for this project come from [this](http://groupware.les.inf.puc-rio.br/har) source

###Loading Libraries and Data
```{r}
library(caret)
library(randomForest)
library(e1071)
# load data files

    #test data
    pmltest <- read.csv("./pml-testing.csv")

    #train data
    pmltrain <- read.csv("./pml-training.csv")
```

###Processing
The data will now be be split into two sets. training and testing so that cross validation can be performed.

```{r}
library(caret)
library(ggplot2)
library(lattice)
library(kernlab)
library(randomForest)
#str(pmltrain)
#str(pmltest)

trainidx <- createDataPartition(pmltrain$classe,p=.9,list=FALSE)
traindata = pmltrain[trainidx,]
testdata = pmltrain[-trainidx,]
set.seed(32768)
nzv <- nearZeroVar(traindata)
trainnzv <- traindata[-nzv]
testnzv <- testdata[-nzv]
pmltestnzv <- pmltest[-nzv]

ftridx <- which(lapply(trainnzv,class) %in% c('numeric'))
trainnzv1 <- preProcess(trainnzv[,ftridx], method=c('knnImpute'))
```

###Prediction
A prediction will now be made using Random Forest

```{r}
pred1 <- predict(trainnzv1, trainnzv[,ftridx])
predtrain <- cbind(trainnzv$classe,pred1)
names(predtrain)[1] <- 'classe'
predtrain[is.na(predtrain)] <- 0

pred2 <- predict(trainnzv1, testnzv[,ftridx])
predtest <- cbind(testnzv$classe, pred2)
names(predtest)[1] <- 'classe'
predtest[is.na(predtest)] <- 0

predpmltest <- predict(trainnzv1,pmltestnzv[,ftridx] )
```

###Modeling
```{r}
model <- randomForest(classe~.,data=predtrain)

predtrain1 <- predict(model, predtrain) 
print(table(predtrain1, predtrain$classe))

training <- as.data.frame(table(predtrain1, predtrain$classe))

predtest1 <- predict(model, predtest) 
print(table(predtest1, predtest$classe))
```

###Results
```{r}
predanswers <- predict(model, predpmltest) 
predanswers

pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_write_files(as.character(predanswers))
```
